\pkg{pksensi}: An R package to apply sensitivity analysis in pharmacokinetic models

### 2019-01-31 manuscript/manuscript.Rmd
update

[@R-sensitivity]

[@R-sensitivity]

Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs. An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions (Ferretti et al., 2016). Also, this technique has been applied in pharmacology and toxicology research (Loizou et al., 2015; McNally et al., 2012; Scherholz et al., 2018). Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time. The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g. compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants (McNally et al., 2011). It can be further applied to parameter prioritization and parameter fixing before model calibration (Hsieh et al., 2018).

In our previous work (Hsieh et al., 2018), we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim (Bois, 2009), an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of global sensitivity analysis algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

```{r, include=FALSE}
knitr::write_bib(c('sensitivity'))
```

>>>>

Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs. An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions (Ferretti et al., 2016). Also, this technique has been applied in pharmacology and toxicology research (Loizou et al., 2015; McNally et al., 2012; Scherholz et al., 2018). Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time. The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g. compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants (McNally et al., 2011). It can be further applied to parameter prioritization and parameter fixing before model calibration [@fphar201800588].

In our previous work [@fphar201800588], we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim (Bois, 2009), an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of global sensitivity analysis algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

```{r, include=FALSE}
knitr::write_bib(c('sensitivity','FME', 'multisensi', 'ODEsensitivity'))
```
